Refinements for open automata

### 2021-12-09 2021-Refinement/paper/refinementarticle.tex
theorem 2 and proof sketch

\TODO{say explicitly what is the difference in the formalisation compared to previous def}
\begin{definition}[Open automata refinement]\label{Def:OA-Refinement}
For two open automata \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation $R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \\[1pt]
		\nwedge V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj} \\
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] $R$ is deadlock reducing.

@@

\begin{lemma} The relation \(\wrel{}{}{H}\) is a preorder thus enables stepwise refinement.
\end{lemma}
The relation  \(\wrel{}{}{H}\) is reflexive,  \(\wrel{A}{A}{H}\),  by taking $R\mpar{s_1, s_2} \mapsto \displaystyle\bigwedge_\subbox{v \in vars(s_1)} {v=v}$ to be $s_1= s_2$  and checking the above conditions (Definition \ref{Def:OA-Refinement}), we can see that\\

\begin{theorem}[Transitivity]
If $\wrel{A_1}{A_2}{H}$ and $\wrel{A_2}{A_3}{H}'$, then $\wrel{A_1}{A_3}{H\cap H'}$.
\end{theorem}
\TODO{proof in appendix + proof sketch}

@@

\section{a recuperer si on se rend comtpte que c'est utile}
When we will introduce refinements in Section \ref{sec:prelref}, setting an undefined variable will be considered a valid refinement, for instance a \(5\) bits register is a particular \(n\) bits register.

\appendix
\section{Proof of transitivity}
\begin{quote}
If $\wrel{A_1}{A_2}{H}$ and $\wrel{A_2}{A_3}{H'}$, then $\wrel{A_1}{A_3}{H\cap H'}$.
\end{quote}
\proof 
If $\wrel{A_1}{A_2}{H}$ then there is $R_{12}$ a relation between states
of $A_1$ and of $A_2$;  If $\wrel{A_2}{A_3}{H'}$ then there is $R_{23}$ a relation between states of $A_2$ and of $A_3$. We build a relation between
 states of $A_1$ and of $A_3$ as follows:  for each pair of states $s_1$, $s_3$, for each state $s_2$ such that $R_{12}$ relates $s_1$ and $s_2$, and $R_{23}$ relates $s_2$ and $s_3$.
Let $R_{13}$ be the relation:\\
  \[R_{13}(s_1,s_3)=\bigvee_{s_2\in S_2}\mpar{R_{12}(s_1,s_2) \land R_{23}(s_2,s_3) } \]

@@

In the same way as this formula is valid for all $s_{2p} \in (s_{2p})^{p\in P}$, we get:

@@

>>>>

\TODO{say explicitly what is the difference in the formalisation compared to previous def}
\begin{definition}[Open automata refinement]\label{Def:OA-Refinement}
For two open automata\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation $R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \\[1pt]
		\nwedge V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj} \\
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] $R$ is deadlock reducing.

@@

\begin{lemma} The relation \(\wrel{}{}{H}\) is a preorder thus enables stepwise refinement.
\end{lemma}
The relation  \(\wrel{}{}{H}\) is reflexive,  \(\wrel{A}{A}{H}\),  by taking $R\mpar{s_1, s_2} \mapsto \displaystyle\bigwedge_\subbox{v \in vars(s_1)} {v=v}$ to be $s_1= s_2$  and checking the above conditions (Definition \ref{Def:OA-Refinement}), we can see that \TODO{finish the sentence}\\

\begin{theorem}[Transitivity]
If $\wrel{A_1}{A_2}{H}$ and $\wrel{A_2}{A_3}{H'}$, then $\wrel{A_1}{A_3}{H\cap H'}$.
\end{theorem}

Appendix~\ref{sec:proof-transitivity} presents the proof. It is done classically by identifying the relation between $A_1$ and $A_3$ that is a refinement. What is less classical is the definition of this relation because it is a boolean formula. For each couple of states  $s_1$ and $s_3$ of $A_1$ and $A_3$ we build a a formula that defines the refinement relation. To do this, we take the disjunction of formulas relating $s_1$ and $s_3$, and passing by all states $s_2$ of $A_2$. More precisely, we define a relation of the following form:
  \[R_{13}(s_1,s_3)=\bigvee_{s_2\in S_2}\mpar{R_{12}(s_1,s_2) \land R_{23}(s_2,s_3) } \]
We then prove that this relation defines is a refinement, according to Definition~\ref{Def:OA-Refinement}.

\TODO{find an informal introduction to the following:}
The composition of $A_1$ and $A_3$ is a refinement of the composition of $A_2$ and $A_3$ provided $A_1$ is a refinement of $A_2$ and the hole in which we are composing inside $A_1$ and $A_2$ is tracked by the refinement relation.

\begin{theorem}[Context refinement]
Let $A_1$, $A_2$ and $A_3$ be three open automata with $\wrel{A_1}{A_2}{H}$. 
Suppose that \(k \in H\) and that \(A_1\subst{A_3}{k}\) is non-locking.
We have: \[\wrel{A_1\subst{A_3}{k}}{A_2\subst{A_3}{k}}{J_3 \uplus H \setminus \mbrc{k}}\].
\end{theorem}

@@

\section{a recuperer si on se rend compte que c'est utile}
When we will introduce refinements in Section \ref{sec:prelref}, setting an undefined variable will be considered a valid refinement, for instance a \(5\) bits register is a particular \(n\) bits register.

\appendix
\section{Proof of Transitivity}\label{sec:proof-transitivity}
\begin{quote}
If $\wrel{A_1}{A_2}{H}$ and $\wrel{A_2}{A_3}{H'}$, then $\wrel{A_1}{A_3}{H\cap H'}$.
\end{quote}
\proof 
If $\wrel{A_1}{A_2}{H}$ then there is $R_{12}$ a relation between states
of $A_1$ and of $A_2$;  If $\wrel{A_2}{A_3}{H'}$ then there is $R_{23}$ a relation between states of $A_2$ and of $A_3$. We build a relation between
 states of $A_1$ and of $A_3$ as follows:  for each pair of states $s_1$, $s_3$, for each state $s_2$ such that $R_{12}$ relates $s_1$ and $s_2$, and $R_{23}$ relates $s_2$ and $s_3$.
Let $R_{13}$ be the relation:\\
  \[R_{13}(s_1,s_3)=\bigvee_{s_2\in S_2}\mpar{R_{12}(s_1,s_2) \land R_{23}(s_2,s_3) } \]

@@

This formula is valid for all $s_{2p} \in (s_{2p})^{p\in P}$,  thus we can state the following:

@@

\section{Proof of Context Refinement}
